Quantum Walks and Application to Quantum Money

### 2025-04-08 draft.tex
chapters 5,6 added

\documentclass[12pt]{report}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{physics}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{setspace}

@@

@@

\section{Hamiltonian Structure and Sparsity}

\section{Quantum Walk-Based and QSP Simulation}

\section{Complexity Analysis}

@@

@@

>>>>

\documentclass[12pt]{report}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{physics}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\newcommand{\cas}{\mathrm{cas}}
\newcommand{\cht}{\mathsf{H}}
\newcommand{\qht}{\mathsf{QHT}}
\newcommand{\cft}{\mathsf{F}}
\newcommand{\qft}{\mathsf{QFT}}
\newcommand{\comph}{\mathsf{cmpIndex}}
\newcommand{\gen}{\mathsf{Gen}}
\newcommand{\ver}{\mathsf{Ver}}

@@

% \section{Hamiltonian Structure and Sparsity}

% \section{Quantum Walk-Based and QSP Simulation}

% \section{Complexity Analysis}

@@

We now present our public-key quantum money scheme based on group actions and the Hartley transform. We begin by reviewing Zhandry’s quantum money scheme from abelian group actions (which uses the Fourier transform) as a baseline. Then we describe our Hartley-transform-based variant and discuss its properties. The quantum Hartley transform will play a central role in both the minting and verification of banknotes.

\section{Public-Key Quantum Money: Definitions}
A public-key quantum money scheme consists of two algorithms, traditionally called $\mathsf{Mint}$ and $\mathsf{Verify}$:
\begin{itemize}
    \item $\mathsf{Gen}(1^n)$ is a probabilistic quantum polynomial-time (QPT) algorithm that, given a security parameter $n$, outputs a pair $(s, |\$\rangle)$. Here $s$ is a classical string called the \emph{serial number} (or public key) and $|\$\rangle$ is a quantum state representing the banknote. We often call the pair $(s, |\$\rangle)$ a banknote, and $s$ by itself the serial number of that banknote.
    \item $\mathsf{Verify}(s, |\$\rangle)$ is a QPT algorithm (which may be implemented as a quantum circuit acting on the state $|\$\rangle$ and some ancillary workspace). It takes as input a serial number $s$ and an alleged quantum banknote state $|\$\rangle$. It outputs $\mathtt{accept}$ (valid) or $\mathtt{reject}$ (invalid).
\end{itemize}

We require two properties from such a scheme:
\begin{description}
    \item[Correctness:] For any valid output $(s, |\$\rangle)$ of $\mathsf{Mint}(1^n)$, the verification algorithm accepts with high probability. Formally, $\Pr[\mathsf{Verify}(s, |\$\rangle) = \mathtt{accept}]$ should be very close to 1 (typically we require it to be $1 - negl(n)$, where $negl(n)$ denotes a negligible function in $n$).
    \item[Security (Unforgeability):] No efficient adversary, given one or more genuine banknotes, can produce a new valid banknote. In the strongest formulation, if an adversary is given a single valid $(s, |\$\rangle)$, it should be infeasible for them to produce two quantum states $|\$'\rangle$ and $|\$''\rangle$ such that $\mathsf{Verify}(s, |\$'\rangle) = \mathsf{Verify}(s, |\$''\rangle) = \mathtt{accept}$ (i.e., they cannot create a second copy that passes verification while retaining the original). More generally, even given multiple banknotes, the adversary should not be able to create new ones with fresh serial numbers that pass verification. This captures the idea that quantum money cannot be counterfeited.
\end{description}

In a \emph{public-key} scheme, the verifier uses only the public information (here, the serial number and any scheme-wide public parameters) to check validity, and does not need any secret key. The bank (mint) may have a secret key to produce the states, but after minting, the bank’s involvement is not needed. In our scheme, the serial number will effectively contain the information needed for public verification (similar to how a classical bill might have a number anyone can look up, except here it’s used in a quantum procedure).

We next outline Zhandry’s scheme, which our scheme builds upon.

\section{Fourier-Based Quantum Money from Group Actions}
Zhandry’s quantum money scheme [31] is built on an abelian group action $G \curvearrowright X$. We have a family of group actions parameterized by the security parameter $n$ (so the size of $G$ grows with $n$). For simplicity, assume $|G| = N$. The scheme also fixes a particular element $x_0 \in X$ that everyone knows (a public base point in $X$).

The minting algorithm $\mathsf{Mint}$ in Zhandry’s scheme works as follows (informally):
\begin{enumerate}
    \item On input $1^n$, sample a random group element $g \leftarrow G$. This $g$ will serve as the serial number $s$ for the banknote.
    \item Prepare two registers: the first is a group element register, initialized to $|0_G\rangle$ (an encoding of the identity element of $G$), and the second is an $X$-register initialized to $|x_0\rangle$.
    \item Apply the quantum Fourier transform over $G$ to the first register. This creates a uniform superposition over $G$: 
    \[
    \frac{1}{\sqrt{N}} \sum_{h \in G} |h\rangle \otimes |x_0\rangle\,.
    \] 
    \item Now apply the group action in a controlled manner: for each basis state $|h\rangle$ in the first register, apply $h$’s action on the second register. This transforms the joint state into 
    \[
    \frac{1}{\sqrt{N}} \sum_{h \in G} |h\rangle \otimes |h * x_0\rangle\,,
    \] 
    which is an entangled state between a superposition of group elements and their corresponding $X$-elements.
    \item Apply the inverse QFT (which is the adjoint of the QFT) on the first register. This recombines the amplitudes in the first register. The result (by properties of the QFT) can be shown to be 
    \[
    \frac{1}{\sqrt{N}} \sum_{h \in G} \omega^{\langle g,\,h \rangle} |h\rangle \otimes |h * x_0\rangle\,,
    \] 
    where $\omega^{\langle g,\,h \rangle}$ denotes the phase picked up that depends on $g$ and $h$. In fact, this state can be written as $|g\rangle \otimes \frac{1}{\sqrt{N}}\sum_{h \in G} |h * x_0\rangle$; essentially, the first register ends up in $|g\rangle$ (the serial number) and the second register is an equal superposition of all elements of $X$ in the orbit of $x_0$ (which is $X$ itself, since the action is transitive) but weighted by a phase related to $g$. This is the so-called group-action Fourier state.
    \item The first register (which contains $|g\rangle$) is measured in the computational basis, yielding the random serial number $g$ and collapsing the second register to a state that we will denote $|\$_g\rangle$. This $|\$_g\rangle$ is the actual quantum money state corresponding to serial $g$.
    \item Output the pair $(s = g,\; |\$\rangle = |\$_g\rangle)$.
\end{enumerate}

In the above, a bit of algebra reveals that $|\$_g\rangle = \frac{1}{\sqrt{N}} \sum_{h \in G} \chi_g(h)\, |h * x_0\rangle$, where $\chi_g(h) = \omega^{\langle g,\,h \rangle}$ is essentially the character of $G$ indexed by $g$. In other words, $|\$_g\rangle$ is (up to global phase) the state one gets by applying the representation (character) $\chi_g$ to weight the superposition over the orbit $X$. These states for different $g$ are almost orthogonal and can be distinguished with the right measurement.

The verification algorithm $\mathsf{Verify}(g, |\$_g\rangle)$ for Zhandry’s scheme goes as follows:
\begin{enumerate}
    \item Given the purported serial number $g \in G$ and an input state $|\psi\rangle$ (which should equal $|\$_g\rangle$ if genuine), attach an auxiliary register prepared in $|x_0\rangle$.
    \item Perform a controlled group action “kickback” using $g$ on the auxiliary register. Specifically, perform the unitary that maps $|y\rangle \otimes |x_0\rangle \mapsto |y\rangle \otimes |g * x_0 \rangle$ for each basis element $|y\rangle$ of the first register. (This operation uses the classical $g$ to act on the second register conditioned on the first—essentially, it multiplies the second register’s state by $g$ if the first register is in an appropriate basis state. There are various ways to implement this, but conceptually it’s a controlled operation using the known $g$.)
    \item Now measure the second register (the auxiliary one) in the computational basis of $X$ to see if it is still $|x_0\rangle$. If the state $|\psi\rangle$ was the correct Fourier state $|\$_g\rangle$, one can show that this measurement will yield $x_0$ with high probability (constructive interference causes the $x_0$ component to be strong). If $|\psi\rangle$ was not the correct state, the measurement is unlikely to give $x_0$.
    \item Accept if and only if the measurement outcome is $x_0$ (i.e. the second register returns to the base state).
\end{enumerate}

The intuition is that the genuine state $|\$_g\rangle$ contains a phase that exactly cancels out the action of $g$ in the verification step, causing the auxiliary register to end up back at $x_0$. A counterfeit state, by contrast, would not generally have the right phase relationship, and the auxiliary register’s state would be some superposition over $X$ that is not concentrated on $x_0$, hence likely giving a different outcome upon measurement.

Zhandry’s scheme is shown to be correct (valid notes pass) and is believed to be secure under the assumption that the group action is one-way. The hardness essentially boils down to: given $x_0$ and $|\$_g\rangle$, an adversary cannot figure out a different $g'$ that would produce a second valid state $|\$_{g'}\rangle$ or replicate $|\$_g\rangle$ (because that would solve the hidden subgroup or discrete log type problem in the group action).

\section{Hartley-Transform Money Scheme Construction}
Our scheme modifies the above Fourier-based construction by utilizing the \textbf{quantum Hartley transform (QHT)} in place of the QFT. Since the Hartley transform is closely related to the Fourier transform (essentially it produces the real and imaginary parts in a single real-valued transform), one might expect this substitution to be straightforward. And indeed, for the minting procedure it largely is: we will use the QHT to create real-amplitude superposition states.

Concretely, assume again a group action $G \curvearrowright X$ with $|G|=N$. We fix the same generating set $S$ and basepoint $x_0$. The $\mathsf{Mint}_{H}$ (Hartley mint) algorithm:
\begin{enumerate}
    \item Sample a random $g \leftarrow G$ to be the serial number.
    \item Prepare $|0_G, x_0\rangle$ as before.
    \item Apply the \emph{quantum Hartley transform} over $G$ to the first register. The Hartley transform (on an abelian group of order $N$) is defined as:
    \[
       \mathrm{QHT}: |h\rangle \mapsto \frac{1}{\sqrt{N}} \sum_{k \in G} \cas \!\Big(\frac{2\pi \langle k,h \rangle}{N}\Big) |k\rangle,
    \] 
    where $\mathrm{cas}(\theta) = \cos \theta + \sin \theta$ and $\langle k,h \rangle$ denotes some pairing like an exponent or dot product (for cyclic groups this is just $kh$). In simpler terms, $\mathrm{QHT}(|0_G\rangle) = \frac{1}{\sqrt{N}}\sum_{k \in G} |k\rangle$ (because $cas(0)=1$), \emph{the same uniform superposition as the QFT produces for $|0\rangle$}. So after this step, we again have $\frac{1}{\sqrt{N}}\sum_{h \in G} |h\rangle|x_0\rangle$, but crucially the coefficients are all \emph{real} and positive (in fact all $1/\sqrt{N}$).
    \item Perform the controlled group action by the first register on the second: $\frac{1}{\sqrt{N}}\sum_{h} |h\rangle |h*x_0\rangle$.
    \item Apply the inverse QHT on the first register (which being its own inverse up to a constant factor, is the same operation again since Hartley is symmetric like Fourier). Now here is where things differ: after this operation, the joint state will be 
    \[
       \frac{1}{\sqrt{N}} \sum_{h \in G} \cas\!\Big(\frac{2\pi \langle g,h \rangle}{N}\Big)\, |h\rangle \otimes |h*x_0\rangle.
    \] 
    Because the Hartley transform combines what would be cosine and sine terms, this state is not as nicely factorable as in the Fourier case. In the Fourier case, we got $|g\rangle \otimes \text{(orbit state)}$. In the Hartley case, we do not get a single basis state in the first register; instead we get a superposition in the first register with real coefficients $c_h = cas(2\pi\langle g,h\rangle/N)$. However, note that $c_h$ is just some real number in $[-1,1]$ depending on $g$ and $h$. Another way to express the above state is:
    \[
      |\$\rangle = \frac{1}{\sqrt{N}} \sum_{x \in X} \Big(\frac{1}{\sqrt{N}}\sum_{h: h*x_0 = x} \cas\!\big(\frac{2\pi \langle g,h \rangle}{N}\big)\Big)\, |x\rangle.
    \] 
    Since for a given $x\in X$, all $h$ such that $h*x_0=x$ have the form $h = u g$ for a unique $u \in G$ (because the action is free and transitive), one can simplify the internal sum using properties of the cas function. But we need not dive into that here.
    \item No measurement of the first register is performed. In fact, we do not have a nice product state separation, so we cannot simply measure one register to get $g$. Instead, the output of minting is just the quantum state $|\$\rangle$ (which is entangled between what used to be two registers) and the known classical serial number $g$ (we know what $g$ we picked).
\end{enumerate}

The resulting money state $|\$\rangle$ has only real amplitudes (because the initial superposition was real and the controlled operation doesn’t introduce complex numbers, and the Hartley transform yields real coefficients). In particular, in the computational basis of $X$, the state $|\$\rangle$ is some real-vector state (with both positive and negative values possible). We can think of it as a “Hartley orbit state” corresponding to $g$.

The verification algorithm in our Hartley-based scheme attempts to mirror the Fourier-based verification: it will use the given serial $g$ and try to check the state. A first idea would be: do exactly the same as Zhandry’s verify (the group-action kickback and measure). However, this fails in certain cases. Specifically, the Hartley transform being real causes an ambiguity: it turns out that there exist distinct $g \neq g'$ for which the Hartley-based states $|\$\rangle_g$ and $|\$\rangle_{g'}$ are not orthogonal, and the original verification might accept a forged state that is a certain superposition. In fact, the verification algorithm might accept an illegitimate state with some probability bounded away from zero. This is the “breakdown” of the straightforward Hartley substitution that we alluded to earlier.

To address this, our verification algorithm takes a more sophisticated approach:
\begin{enumerate}
    \item Given $(g, |\psi\rangle)$, it first uses the \textbf{quantum walk-based phase estimation} method (to be detailed in Chapter 6) to \emph{compute the eigenphase corresponding to $g$}. In short, we will treat $|\psi\rangle$ as an eigenvector of the group action adjacency matrix with a certain eigenvalue $\lambda$ related to $g$. We run a procedure to estimate $\lambda$ to high precision.
    \item From the estimated eigenvalue(s), we reconstruct a candidate group element $g^*$ that we believe generated the state. (This uses the fact that by running phase estimation for multiple carefully chosen walk durations, one can solve for $g$ as shown in [31].)
    \item We check if $g^* = g$ (the serial number provided). If not, we reject, since the state does not match the serial. If yes, we accept.
\end{enumerate}

This procedure essentially extracts the hidden group element from the state and compares it to the claimed serial. If an adversary tries to forge a state $|\psi\rangle$ for a known serial $g$, the only way to consistently pass verification is if $|\psi\rangle$ truly encodes the same $g$ in its eigen-spectrum. A fake note that was causing trouble for the naive Hartley verification (e.g., a state that is a mixture of two different “Hartley orbit” states) would yield an eigenphase that does not correspond to a single valid $g$, and the algorithm would detect the mismatch.

The key quantum tool enabling this verification is the continuous-time quantum walk on the group action graph and the ability to simulate it, which we prepared in Chapter 4. The walk Hamiltonian $A(X,S)$ has the genuine money state $|\$\rangle_g$ as an eigenstate with a known eigenvalue $\lambda(g)$. By running phase estimation on $e^{-iA t}$ with the state $|\psi\rangle$, we attempt to measure that eigenvalue. We repeat for a few different values of $t$ (the “twists” or different phases) to get enough information to pinpoint $g$. This approach can be viewed as adding extra “twists” to the verification: effectively, each choice of $t$ in phase estimation is like looking at the state in a different interference fringe pattern, analogous to applying different group actions as twists in the verification procedure. This is why we described the new verification as using group action twists—by varying a continuous parameter related to the group action’s eigenvalues, we break the symmetry that allowed a counterfeit to slip through.

We defer the full analysis of the success probabilities and how many repetitions are needed to Chapter 6, where we describe the verification algorithm in detail. For now, the takeaway is that by using the quantum walk approach, we can reliably verify Hartley-based money states, restoring security while still enjoying the properties of real amplitudes.

\section{Efficient Quantum Hartley Transform Implementation}
Before moving on to the verification algorithm, we briefly discuss how we implement the quantum Hartley transform (QHT) efficiently, as this is an important practical aspect of our scheme. The Fourier transform is well-known to have efficient circuits; for the Hartley transform, fewer results are available, so we contributed a new circuit construction.

The QHT we need is over an abelian group, which for concreteness one can think of as $\mathbb{Z}_N$ (the general finite abelian group case can usually be reduced to a direct sum of cyclic groups, applying QHT on each cyclic component). Classical Hartley transforms have a recursive structure similar to fast Fourier transforms. In fact, there are known formulas to express a Hartley transform of size $N$ in terms of two Hartley transforms of size $N/2$ plus some additional linear operations [25]. We leveraged such structures to design a divide-and-conquer quantum circuit.

Our algorithm for QHT works roughly as follows:
- If $N$ is even, we can express the length-$N$ Hartley transform in terms of two length-$N/2$ Hartley transforms plus a combination of cheap operations (additions, permutations of data, etc.). We implement this decomposition recursively as a quantum circuit. The base of the recursion is when the size is small (like $N=2$ or $4$, where the transform can be done with a constant number of gates).
- We found that by carefully optimizing this recursion and using some known quantum subroutines for certain linear combinations, the overall gate count is improved compared to naive methods or previous proposals [21, 22]. In particular, our circuit avoids introducing any quantum Fourier transform internally (unlike [22] which computed Hartley via Fourier plus some adjustments), and this yields a purely real rotation-based circuit.

%%%%%%%%%%% az inja %%%%%%%%%%%%%

Our algorithm for the quantum Hartley transform, \( \qht_N \), is modeled after the recursive structure of the quantum Fourier transform algorithm \( \qft_N \). To motivate this, we begin by reviewing the standard approach for computing \( \qft_N \). For clarity, we assume \( N = 2^n \), allowing each element of \( \mathbb{Z}_N \) to be represented using exactly \( n \) qubits. The same recursive framework can be extended to handle general values of \( N \). For any \( a \in \mathbb{Z}_N \), the quantum Fourier transform is defined as:

\begin{align*}
    \qft_N\ket{a}
    & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \omega_N^{ay}\ket{y} \\
    & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \ket{y} + (-1)^a \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \ket{y + N/2} \\
    & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \frac{1}{\sqrt{2}} (\ket{0} + (-1)^a \ket{1}) \ket{y}, %\numberthis\label{eq:qft_alt}
\end{align*}

In the final expression, we have isolated the first qubit for clarity. Let \( \ket{a} = \ket{t}\ket{b} \), where \( b \) denotes the least significant bit of \( a \), implying that \( a = 2t + b \) for some \( t \in \mathbb{Z}_{N / 2} \). Applying \( \qft_{N / 2} \) to the first register yields the state:

\[ \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{2ty} \ket{y} \ket{b}. \]
Next, we apply the phase unitary \( P(y, b) \), defined by \( \ket{y} \ket{b} \mapsto \omega_N^{by} \ket{y} \ket{b} \), followed by a Hadamard transform on the final qubit. This results in the state given in Equation~\eqref{eq:qft_alt}.

We now describe our algorithm for efficiently computing the quantum Hartley transform \( \qht_N \). The key idea is to leverage the recursive structure of \( \qht_N \), analogous to the strategy used for \( \qft_N \). To facilitate this, we begin by rewriting the summation in Equation~\eqref{eq:qht-N} to expose its recursive form. Our approach is as follows:
\begin{equation}
    \label{eq:cas-expand}
	\frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \ket{y}
    = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \ket{y} + \frac{1}{\sqrt{N}} \sum_{y = N / 2}^{N - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \ket{y}.
\end{equation}

The second summation on the right-hand side can be expressed as:
\begin{align*}
	\sum_{y = N / 2}^{N - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \ket{y}
    & = \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi a y}{N} + \pi a \Big) \ket{y + N/2} \\
    & = (-1)^a \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \ket{y + N/2},
\end{align*}
The second equality follows from the angle-sum identity for the \( \cas \) function,
\[
\cas(\alpha + \beta) = \cos(\alpha)\cas(\beta) + \sin(\alpha)\cas(-\beta),
\]
along with the fact that \( \cos(\pi a) = (-1)^a \) for all integers \( a \). Substituting this into Equation~\eqref{eq:cas-expand} yields:
\begin{align}
	\frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \ket{y}
    & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi a y}{N} \Big) (\ket{y} + (-1)^a \ket{y + N/2}) \nonumber \\
    & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi a y}{N} \Big) \frac{1}{\sqrt{2}} (\ket{0} + (-1)^a \ket{1}) \ket{y}, \label{eq:qht-alt} 
\end{align}
In the final equality, the most significant qubit has been isolated to emphasize its contribution to the overall transformation.

We now demonstrate how to compute \( \qht_N \) recursively. For a given \( a \in \mathbb{Z}_N \), we decompose the basis state as \( \ket{a} = \ket{t}\ket{b} \), where \( b \) denotes the least significant bit, such that \( a = 2t + b \) for some \( t \in \mathbb{Z}_{N/2} \). Assuming the existence of an efficient quantum circuit for computing \( \qht_{N/2} \), we proceed to construct \( \qht_N \) in a recursive fashion. To begin the transformation on \( \ket{a} \), we introduce an ancilla qubit initialized to \( \ket{0} \), yielding the joint state \( \ket{0}\ket{t}\ket{b} \). We then proceed as follows:
\begin{align*}
	\ket{0}\ket{t}\ket{b}
    & \mapsto \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi t y}{N / 2} \Big) \ket{0} \ket{y} \ket{b} & (\mathbf{1} \otimes \qht_{N / 2} \otimes \mathbf{1}) \\
    & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) \ket{0} \ket{y} \ket{b} \\
    & \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) (\ket{0} + \ket{1}) \ket{y}\ket{b}. & (H \otimes \mathbf{1})
\end{align*}
Next, we apply a controlled negation operation, where the ancilla qubit is flipped if and only if the least significant bit \( b \) is equal to 1.
\[
    V: \ket{0}\ket{y} \mapsto \ket{0}\ket{y}, \quad \ket{1}\ket{y} \mapsto \ket{1}\ket{N / 2 -y},
\]
\[
    \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) \ket{0} \ket{y} \ket{b} + \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) \ket{1} \ket{-y} \ket{b}.
\]
By performing a change of variables in the second summation and using the identity \( \cas\left(\frac{4\pi t (N / 2 - y)}{N}\right) = \cas\left(-\frac{4\pi t y}{N}\right) \), we obtain the state:
\[
    \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \Big( \cas\Big( \frac{4\pi t y}{N} \Big) \ket{0} + \cas\Big( -\frac{4\pi t y}{N} \Big) \ket{1} \Big) \ket{y} \ket{b}.
\]

We define the single-qubit rotation \( R(\theta) \) as the unitary operator acting on a single qubit given by:
\begin{equation}
    \label{eq:sine-rot}
    R(y, b) = 
    \begin{bmatrix}
        \cos(2\pi b y / N) & \sin(2\pi b y / N) \\
        -\sin(2\pi b y / N) & \cos(2\pi b y / N)
    \end{bmatrix},
\end{equation}
Now, consider the unitary \( U_R \) defined by
\[
U_R : \ket{c}\ket{y}\ket{b} \mapsto \left(R(y, b)\ket{c}\right)\ket{y}\ket{b},
\]
where \( R(y, b) \) is a single-qubit rotation depending on \( y \) and \( b \). Applying \( U_R \) followed by the controlled negation \( V \), we obtain the state:
\begin{align*}
    \ket{\phi_1}
    & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \Big( \cas\Big( \frac{2\pi ay}{N} \Big) \ket{0} + \cas\Big(-\frac{2\pi a(N /2 - y)}{N} \Big) \ket{1} \Big) \ket{y} \ket{b} \\
    & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi ay}{N} \Big) \Big( \ket{0} + (-1)^b \ket{1} \Big) \ket{y} \ket{b},
\end{align*}
Here, we have used the identity \( \cas\left(-\pi a + \frac{2\pi a y}{N}\right) = (-1)^a \cas\left(\frac{2\pi a y}{N}\right) \), along with the fact that \( (-1)^a = (-1)^b \). Applying the Hadamard transform to the first qubit yields:
\[
    \ket{\psi} = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi ay}{N} \Big) \ket{b} \ket{y} \ket{b}
\]
Next, we uncompute the first qubit by applying a \textsc{CNOT} gate with the last qubit as control. We then apply a Hadamard transform followed by a swap operation, resulting in the state:
\begin{align*}
    \ket{\psi}
    & \mapsto \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi ay}{N} \Big) \ket{0} \ket{y} \ket{b} & (\textsc{cnot}) \\
    & \mapsto \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi ay}{N} \Big) \ket{0} \ket{y} \frac{1}{\sqrt{2}}(\ket{0} + (-1)^b \ket{1}) & (\mathbf{1} \otimes H) \\
    & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi ay}{N} \Big) \ket{0} \ket{y} \frac{1}{\sqrt{2}}(\ket{0} + (-1)^a \ket{1}). & ((-1)^b = (-1)^a) \\
\end{align*}
The final summation matches precisely with Equation~\eqref{eq:qht-alt}, which defines the quantum Hartley transform of \( \ket{a} \). Thus, we have successfully implemented the transformation:
\[
\ket{a} \mapsto \qht_N \ket{a}.
\]
\[ \ket{0}\ket{a} \mapsto \ket{0} \qht_N\ket{a}. \]
The complete procedure is summarized in the following algorithm.

% \begin{algorithm}[$\qht_N$] \
\paragraph{algorithm $\qht_N$}
    % \label{alg:qht-N}
    % \begin{description}[font = \normalfont\itshape, itemsep = 0mm, parsep = 0mm, topsep = 1mm]
    %     \item [Input:] quantum state $\ket{\psi} \in \C^N$, where $N = 2^n$
    %     \item [Output:] quantum state $\qht_N\ket{\psi}$
    % \end{description}

\begin{enumerate}%[itemsep = 0mm, parsep = 1mm, topsep = 1mm]
    \item Initialize an ancilla qubit to $0$ to obtain the state $\ket{0}\ket{\psi}$
    \item Compute $\mathbf{1} \otimes \qht_{N / 2} \otimes \mathbf{1}$ recursively.
    \item Apply $H \otimes \mathbf{1}$.
    \item\label{stp:ngt} Apply the controlled negation $\ket{0}\ket{y} \mapsto \ket{0}\ket{y}, \ket{1}\ket{y} \mapsto \ket{1}\ket{N / 2 -y}$ to the first two registers.
    \item Apply the unitary $U_R$.
    \item Apply $H \otimes \mathbf{1}$
    \item Apply \textsc{cnot} to the first and last qubits.
    \item Apply $\mathbf{1} \otimes H$.
    \item Trace out the first qubit
\end{enumerate}
% \end{algorithm}

% \begin{theorem}
\paragraph{Theorem.}
    \label{thm:qht-cost}
    Algorithm \ref{alg:qht-N} is correct and can be implemented using $\approx \log^2 N + O(\log N)$ elementary gates.
% \end{theorem}

% \begin{proof}
\paragraph{proof.}
    The correctness of the algorithm follows from the preceding discussion. Except for the unitary \( U_R \) and the negation unitary in Step~\ref{stp:ngt}, all steps of the algorithm can be implemented using \( O(1) \) elementary gates. The negation operation in Step~\ref{stp:ngt} can be realized using approximately \( \lceil \log N \rceil \) elementary gates.

    To implement the unitary \( U_R \), which requires constructing the conditional rotation operator \( R(y, b) \) for arbitrary \( y \) and \( b \), we make use of the two-qubit controlled rotations:
    \[
    R_j = \ket{0}\bra{0} \otimes \mathbf{1} + \ket{1}\bra{1} \otimes
    \begin{bmatrix}
    \cos\left(\frac{2\pi 2^j}{N}\right) & \sin\left(\frac{2\pi 2^j}{N}\right) \\
    -\sin\left(\frac{2\pi 2^j}{N}\right) & \cos\left(\frac{2\pi 2^j}{N}\right)
    \end{bmatrix},
    \]
    for \( j = 0, 1, \dots, n - 1 \). When \( b = 0 \), we have \( R(y, 0) = \mathbf{1} \), and when \( b = 1 \), the operator \( R(y, 1) \) is the product of those \( R_j \) for which the \( j \)th bit of \( y \) is 1. Therefore, for a binary string \( y \) of length \( k = \lceil \log y \rceil \), we can implement \( U_R \) using at most \( k \) gates from the set \( \{R_j\} \).

    Let \( T(N) \) denote the gate complexity of Algorithm~\ref{alg:qht-N} for an input of dimension \( N \). Assuming access to the \( R_j \) gates as elementary operations, the recurrence relation becomes
    \[
    T(N) \approx T(N / 2) + 2\log N + O(1).
    \]
    Solving this recurrence yields \( T(N) \approx \log^2 N + O(\log N) \), as claimed.
%     \end{proof}
%%%%%%%%%%%  ta inja  %%%%%%%%%%%

- The resulting complexity for an $N$-point QHT is $O(N \log N)$ basic quantum gates (up to polylog factors for precision), which is on par (up to constant factors) with the complexity of a QFT circuit. We also derived explicit constants for small cases to show improvements over prior work.

For the purposes of this thesis, the main point is: \emph{we can implement the Hartley transform needed for our money scheme with polynomial efficiency}. Thus, switching to the Hartley transform does not introduce any prohibitive cost. All steps of minting and verification remain efficient. Having covered the construction of the money scheme and the necessary tools, we now proceed to the final piece: the detailed verification algorithm using quantum walks, and an analysis of how it validates genuine banknotes and foils counterfeit attempts.
\vspace{1em}

@@

The Hartley-based quantum money scheme introduced in the previous chapter requires a new verification procedure to address the issues that arise from using real amplitude states. In this chapter, we describe and analyze the verification algorithm, which leverages continuous-time quantum walks on the group action graph (and their efficient simulation from Chapter 4) to extract the information needed to authenticate a banknote.

    \section{Challenges with Naïve Verification}
    Before detailing the new algorithm, let us briefly recap why the straightforward approach fails. In the Fourier-based scheme, verification was done by a single “kickback” operation using the claimed serial $h$, and measuring an auxiliary register. In the Hartley-based scheme, if we attempted the analogous one-step verification, we would perform the controlled-$h$ action and measure the auxiliary system. A genuine state $|\$_h\rangle$ (the Hartley money state for serial $h$) would cause some interference pattern in the auxiliary register, but unlike the Fourier case, it does not return the auxiliary to $|x_0\rangle$ deterministically. In fact, there is an ambiguity: certain superpositions of eigenstates corresponding to $h$ and $-h$ (or other group-related variants) can produce the same measurement statistics in that one-step test. This means an adversary might prepare a counterfeit state that is not a legitimate $|\$_h\rangle$ but still passes the one-step verification with non-zero probability. Essentially, the Hartley transform being real means we lost some phase information, and a single measurement cannot distinguish some mirrored states.

    To overcome this, our strategy is to perform a more complete measurement of the state’s “phase spectrum.” Instead of just one operation and measurement, we will use the quantum walk (with Hamiltonian $A = A(X,Q)$ as defined earlier) to perform a form of phase estimation.

    \section{Using Quantum Walks to Extract the Serial}

    Given a state of the form $\ket{\mathbb{Z}_N^{(h)} * x}_H$, we demonstrate how the value of $h$ can be extracted using continuous-time quantum walks. For any $u \in \mathbb{Z}_N$, consider the Cayley graph $\Gamma = (\mathbb{Z}_N, E)$ generated by $Q = {\{-u, u}\}$. Let $A$ denote the adjacency matrix of $\Gamma$. The eigenvectors of $A$ are given by $\ket{\mathbb{Z}_N^{(h)} * x}$, with corresponding eigenvalues $\lambda_h = 2\cos(2\pi uh / N)$ for each $h \in \mathbb{Z}_N$. According to Theorem \ref{thm:sparse-sim} and the discussion that follows, the unitary evolution operator $W = e^{iAt}$ can be simulated efficiently to exponential precision. To proceed, we first state the following lemma.
